---
title: "About reproducibility"
author: "inSileco Team"
output:
  xaringan::moon_reader:
    css: [default, rd.css, rd-font.css, "hygge"]
    lib_dir: assets
    seal: false
    nature:
      highlightStyle: dracula
      countIncrementalSlides: false
      beforeInit: "macros.js"
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  dev = "png",
  dpi = 108,
  fig.width = 6,
  fig.height = 4.5,
  fig.align = 'center',
  width = 120
)
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
rfa <- function(...) icon::fontawesome(...)
emj <- function(...) emo::ji(...)
kig <- function(...) knitr::include_graphics(...)
```





---

class: inverse, center, middle

# `r rfa("info-circle")` Reproducible research

![:custom_hr]()

## `r icon::fontawesome("clock")` ~ 30min


???

Comments


---
# Reproducible research `r emj("hand")`

<br>

## Reproducibility

<br>

--

## Reproducibility vs. Replicability

<br>

--

## Reproducibility vs. Replicability vs. Generalizability


---
# Terminology is not clear yet...

- Various terms with different meanings

```{R, echo = FALSE, out.width = "50%"}
kig("img/r_words.png")
```

`r rfa("link")` Goble (2016), [What is Reproducibility?](https://www.slideshare.net/carolegoble/what-is-reproducibility-gobleclean)


---
# Terminology is not clear yet...

## Reproducible vs. Replicable

`r rfa("file")` Barba (2018), [Terminologies for Reproducible Research](https://arxiv.org/abs/1802.03311)

--

### **no distinction** (A)
  
--
  
### **distinction** :
- for some (B1):
  - *reproducible*: authors provide data + computer code to re-create the results.
  - *replicable*: new study same scientific findings as another study, collecting new
  data and new analyses.
--

- for others (B2): opposite definitions

???

similar methods but could be different


---
# Terminology is not clear yet...

## Reproducible vs. Replicable


```{R, echo = FALSE, out.width = "80%"}
kig("img/terminos.png")
```

`r rfa("file")` Barba (2018), [Terminologies for Reproducible Research](https://arxiv.org/abs/1802.03311)


---
# Terminology is being clarified


```{R, echo = FALSE, out.width = "50%"}
kig("img/cover_rrs.png")
```

`r rfa("file")` National Academies of Sciences, Engineering, and Medicine
(2019). [Reproducibility and Replicability in Science](https://www.nap.edu/catalog/25303/reproducibility-and-replicability-in-science). 


---
# Terminology is being clarified

> We define **reproducibility** to mean **computational reproducibility**—­
obtaining consistent computational results using the same **input data**, computational steps, methods, **code**, and conditions of analysis; 

--

> and **replicability** to mean obtaining **consistent results across studies aimed at answering the
same scientific question**, each of which has obtained its own data. 

--

> In short, reproducibility involves the original data and code; replicability involves
new data collection and similar methods used by previous studies. 

`r rfa("file")` National Academies of Sciences, Engineering, and Medicine
(2019). [Reproducibility and Replicability in Science](https://www.nap.edu/catalog/25303/reproducibility-and-replicability-in-science). 

???

1. good practices
2. robustness of a finding science


---
# Terminology is being clarified

> A third concept, **generalizability**, refers to the **extent that results of a study apply** in other contexts or populations that differ from the original one.


`r rfa("file")` National Academies of Sciences, Engineering, and Medicine
(2019). [Reproducibility and Replicability in Science](https://www.nap.edu/catalog/25303/reproducibility-and-replicability-in-science). 

???

what the scope of the finding


---
# Reproducibility is tied to computer science

<br>

### **1. Scientific studies should be replicable (Material & Methods)**

--

<br>

### **2. With the help of technologies scientific studies should now be reproducible (Research compendium)**


???

It's a topic about research but quickly due to its nature we are talking about technologies.


---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/knowledge1.png")
```

??? 

similar with report 

---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/knowledge2.png")
```

---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/knowledge3.png")
```

`r rfa("file")` Goodman et al. (2016). [What does research reproducibility mean?](https://stm.sciencemag.org/content/8/341/341ps12). 


---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/knowledge2.png")
```

???

only retrieve data!!not carry out experiment 
getting more complex in terms of computataional steps.


---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/knowledge4.png")
```


---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/workflow1.png")
```



---
# How do we produce knowledge?

```{R, echo = FALSE, out.width = "55%"}
kig("img/workflow2.png")
```

???

circle : science bringing new tools and with new tools more complex science.

---
# How do we produce knowledge?


## Analyzing data has become increasingly complex 

<br>

--

## - more and more `r rfa("laptop-code")` steps 

<br>

--

## - each step has a certain risk of user error.

--

Recognized in the 80's/90's by [Jon Claerbout](https://en.wikipedia.org/wiki/Jon_Claerbout) and this led to [Redoc](https://ieeexplore.ieee.org/document/881708).


---
# How do we produce knowledge?


## "Publish or perish" system urges scientists to work (too) fast

<br>

--

## and to adopt "questionable research practices" 
  - p-value hacking, 
  - cherry picking,
  - Hypothesizing After Results are Known (HARKing).

`r rfa("file")` Fraser et al. (2018). [Questionable research practices in ecology and evolution](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0200303)

???

also covid money or ego


---
# Why does reproducibility matter?


```{R, echo = FALSE, out.width = "67%"}
kig("img/replic_osf.png")
```

.font80[`r rfa("file")` Open Science Collaboration (2015). [Estimating the reproducibility of psychological science.](https://science.sciencemag.org/content/349/6251/aac4716)]

---
# Why does reproducibility matter?


```{R, echo = FALSE, out.width = "90%"}
kig("img/replic_osf2.png")
```

.font80[`r rfa("file")` Open Science Collaboration (2015). [Estimating the reproducibility of psychological science.](https://science.sciencemag.org/content/349/6251/aac4716)]

???

might very well expect that for ecology

---
# Why does reproducibility matter?


```{R, echo = FALSE, out.width = "65%"}
kig("img/retractation_index.png")
```

.font80[`r rfa("file")` Fang & Casadevall. (2011) [Retracted Science and the Retraction Index.](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3187237/)]



---
# Why does reproducibility matter?

```{R, echo = FALSE, out.width = "55%"}
kig("img/workflow3.png")
```


---
# Why does reproducibility matter?


### .font140[**Can do**: make research findings more robust]

--


### .font140[**Require**: making data and code open]


--

### .font140[**Collateral benefits**]

  - "**Open science**" movement:
    - [Open Science Framework](https://osf.io/) 
    - [rOpenSci](https://ropensci.org/)
    
--
  
  - re-usage of code can lead to an increase of efficiency;
  - improved coding skills;


???

DB and I have developed coding skills for this.

---
# How do reproduce studies


### .font140[**Obstacle**: it takes **a lot** of time]

> It is also important to recognize that there are costs and barriers to shifting to a practice
of reproducible research, particularly when the culture does not recognize the value of
developing this new paradigm or the effort that can be required to develop or learn to use
suitable tools. This is of particular concern to young people who need to earn tenure
or secure a permanent position.


.font80[`r rfa("file")` Stodden et al. (2013).
[Setting the default to reproducible: Reproducibility in Computational and Experimental Mathematics](http://ethics.iit.edu/eelibrary/node/15429)]



???

I've spent a lot of time on making my studies reproducible but not even sure sb re-run my code, not even sure it'll be useful. What is sure is that I didn't spend that time on other paper and I am "more proud of what I have done". But so far, without a system shift, it could be regarded as a virtue. I am very happy to see DFO is taking this very seriously and understood the collateral benefits. Yes it pays off. There is less technical obstacle, there are better practice. You can do more.


---
# A reproducibility spectrum 

```{R, echo = FALSE, out.width = "80%"}
kig("img/spectrum.png")
```

.font80[`r rfa("file")` Peng. (2011).
[Reproducible Research in Computational Science.](https://science.sciencemag.org/content/334/6060/1226)]


--

### Towards full reproducibility, technologies are getting more complex

- entire pipeline in a versioned repository

--

- record all version of software used in the workflow

--

- code run without being impacted by OS or hardware



???

Before making the topics more complex that needed.
We are aiming at given you a sense of this!! 






---

class: inverse, center, middle

# A minimal workflow with `r rfa("r-project")`

![:custom_hr]()

## `r icon::fontawesome("clock")` ~ 20min


???

Questions??






---
# You might not be that far! 

### R is suitable for Reproducible

- open source (no license issues)

--

- strong community actively engaged in making science reproducible

--

- [rOpenSci](https://ropensci.org/)


???

community ==> made of R user 
and so are Python and Julia

---
# Is ?

### YES YES and YES! 

### Ropensci




- input data in one place and never modify them afterwards

- R scripts in one place 

well... that's a very good beginning 


---
# Create a R folder 

Create a meaningful set up 

- R/
- data/
- doc/Rmd <- articulated pipeline 

---
# Create a R folder 

Ensure that all code work from root; ensure form starts 

attached an sessionInfo()


---
# Create a R folder 

Use version comtrol 
Make your in a remote/virtual machine...

towards the gold standard, technologies are increasingly more complex 


---
# IMHO 

- make data avaialble 
- attached local repo in appendix, a zip file
- sessionInfo()

This would be just wonderful.... this would be a huge stepo. 

At the end I'll talk about breofly about a set of R packages and other techno/software that allow omne to ge further,. 


---
# Exercice 

### Description

```{r, echo = FALSE, purl = FALSE}
countdown::countdown(minutes = 15, seconds = 0)
```


---
# Resources `r rfa("link")`

### Reproducibility

- [Barba-group reproducibility syllabus](https://medium.com/hackernoon/barba-group-reproducibility-syllabus-e3757ee635cf#.h603izuwo)
- [*What does research reproducibility mean?*](https://stm.sciencemag.org/content/8/341/341ps12.full)
- [Retracted Science and the Retraction Index](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3187237/)
- [Questionable research practices in ecology and evolution](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0200303)
- [GitHUb thread](https://github.com/ReScience/ReScience-article/issues/5)



---
# Resources `r rfa("link")`

### OSF

- [COS: Strategic Plan, v2.0](https://docs.google.com/document/d/1sqz3appQ73vqa6fP1Gy8KK8HZpotoSGaiJC1XQuvREI/edit)
- [50 cancer studies](https://www.sciencemag.org/news/2018/07/plan-replicate-50-high-impact-cancer-papers-shrinks-just-18)
- [Reproducibility Project: Cancer Biology](https://elifesciences.org/collections/9b1e83d1/reproducibility-project-cancer-biology)


---
# Resources `r rfa("link")`

### Research Compendium 

- [rOpenSci](https://ropensci.org/)
- [R for Reproducible Scientific Analysis](https://swcarpentry.github.io/r-novice-gapminder/)
- [Reproducible Analysis With R](https://nceas.github.io/sasap-training/materials/reproducible_research_in_r_fairbanks/)
- [Touchdown your research with Rmarkdown](https://insileco.github.io/ResearchDown/#1)
- [Formation FRB-Cesab & GDR EcoStat (fr)](https://frbcesab.github.io/datatoolbox/index.html#1)
- [Research Compendium](https://research-compendium.science/) landing page
- [rrrpkg](https://github.com/ropensci/rrrpkg)
- [rrtools](https://github.com/benmarwick/rrtools)




---
# Quote



> The fact that an analysis is reproducible does
not guarantee the quality, correctness, or validity of the published results. (Peng, 2011)


> Unfortunately, the mere reproducibility of
Reproducibility—the ability to recompute
results—and replicability—the chances other
experimenters will achieve a consistent
result—are two foundational characteristics
of successful scientific research. (Leek and Peng 2015)

[Roger D. Peng](https://github.com/rdpeng)
 



---
# Different R version different results


```R
# R 3.6
df <- data.frame(fact = rep(c("F1", "F2"), each = 5), val = runif(10))
plot(val ~ fact, data = df)   # works 
# R 4.0
df <- data.frame(fact = rep(c("F1", "F2"), each = 5), val = runif(10))
plot(val ~ fact, data = df)   # does not work
```



---
# inSileco

<!-- LAST SLIDE -->

- inSileco https://insilecoinc.github.io/
- David Beauchesne [`r rfa("envelope")`](mailto:david.beauchesne@hotmail.com)
- Kevin Cazelles [`r rfa("envelope")`](mailto:kevin.cazelles@gmail.com)

